---
title: "Image Analysis Magenta 2"
author: "Claudia Probst"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 3
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
---
# Experimental Setup

We're investigating the responses of *Arabidopsis thaliana* to varying levels of P-fertilizer supply in presence or absence of single strains of the SynCom9. ARADEEPOPSIS is a software that can be used to study pheotypic traits of *Arabidopsis thaliana* by images. We can investigate biomass, shape and size by different features over time. 

```{r setup, include=FALSE, echo=F, message=F, warning=F}
##clear the objects from memory
rm(list=ls())
#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)
## installs (if necessary) and loads libraries
library(ggplot2)
library(cowplot)
library(tibble)
library(tidyr)
library(ggsignif)
library(datasets)
library(multcompView)
library(dplyr)

#set seed
set.seed(100)

setwd("~/Desktop/MASTER Pflanzenwissenschaften/Experiments/Magenta assay/Magenta single_strain_all")
data<-read.csv(file="aradeepopsis_traits.csv", header=TRUE, sep=",")
str(data)
head(data)  

```

Feautures of interest
```{r foi, echo=F, message=F, warning=F}

# select features of interest
foi <- c("plant_region_red_channel","plant_region_green_channel", "plant_region_blue_channel", "plant_region_chroma_ratio", "plant_region_chroma_difference", "plant_region_chroma_base", "plant_region_green_strength","plant_region_blue_green_ratio", "plant_region_area", "plant_region_filled_area", "plant_region_convex_area","plant_region_equivalent_diameter", "plant_region_major_axis_length","plant_region_minor_axis_length", "plant_region_perimeter","plant_region_eccentricity", "plant_region_extent", "plant_region_solidity", "plant_region_aspect_ratio")
data_foi <- subset.data.frame(data, select = c("file", "assay", "treatment", "date", "day", foi))

rm(data)
```

```{r z_scores, echo=F, message=F, warning=F}
#prepare df
data_foi_norm <- data_foi
data_foi_norm[,(ncol(data_foi_norm)-length(foi)+1):ncol(data_foi_norm)] <- NA

#normalise by z-scores each feature for each time point
assays <- unique(data_foi$assay)

#loop
for (e in assays) {
  for (f in foi) {
    x <- data_foi[data_foi$assay == e, colnames(data_foi) == f] #get values for experiment for foi
    z <- (x-mean(x))/(sd(x)) #compute z_score
    data_foi_norm[data_foi_norm$assay == e, colnames(data_foi_norm) == f] <- z #store it
  }
}

rm(data_foi)
```
STATISTICAL ANALYSIS from the last time point:
We are now looking only at the data from the 05.12.22 (day 21)

```{r final dataset,echo=F, message=F, warning=F}
all_data<-subset(data_foi_norm, day=="day 21")

```


```{r normal distribution last timepoint,echo=F, message=F, warning=F, fig.show="hide"}


LPC<- subset(all_data, treatment=="lowP_control")
LPS<- subset(all_data, treatment=="lowP_SynCom9")
LPF247<- subset(all_data, treatment=="lowP_F247")
LPF112<- subset(all_data, treatment=="lowP_F112")
LPF125<- subset(all_data, treatment=="lowP_F125")
LPF242<- subset(all_data, treatment=="lowP_F242")


#low P control
par(mfrow=c(3:2))
hist(LPC$plant_region_area)
qqplot<-scale(LPC$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPC$plant_region_area)

#low P SynCom9

hist(LPS$plant_region_area)
qqplot<-scale(LPS$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPS$plant_region_area)

#low P F247

hist(LPF247$plant_region_area)
qqplot<-scale(LPF247$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)

shapiro.test(LPF247$plant_region_area)

#low P F112
par(mfrow=c(3:2))
hist(LPF112$plant_region_area)
qqplot<-scale(LPF112$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPF112$plant_region_area)

#low P F125
hist(LPF125$plant_region_area)
qqplot<-scale(LPF125$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPF125$plant_region_area)

#low P F242
hist(LPF242$plant_region_area)
qqplot<-scale(LPF242$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPF242$plant_region_area)



HPC<- subset(all_data, treatment=="highP_control")
HPS<- subset(all_data, treatment=="highP_SynCom9")
HPF247<- subset(all_data, treatment=="highP_F247")
HPF112<- subset(all_data, treatment=="highP_F112")
HPF125<- subset(all_data, treatment=="highP_F125")
HPF242<- subset(all_data, treatment=="highP_F242")


#high P control
par(mfrow=c(3:2))
hist(HPC$plant_region_area)
qqplot<-scale(HPC$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPC$plant_region_area)

#high P SynCom9
hist(HPS$plant_region_area)
qqplot<-scale(HPS$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPS$plant_region_area)

#high P F247
hist(HPF247$plant_region_area)
qqplot<-scale(HPF247$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPF247$plant_region_area)

#high P F112
par(mfrow=c(3:2))
hist(HPF112$plant_region_area)
qqplot<-scale(HPF112$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPF112$plant_region_area)

#high P F125
hist(HPF125$plant_region_area)
qqplot<-scale(HPF125$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPF125$plant_region_area)

#high P F242
hist(HPF242$plant_region_area)
qqplot<-scale(HPF242$plant_region_area)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPF242$plant_region_area)

```

```{r merging subsets, echo=F, message=F, warning=F }

#merging all subsets to one dataframe

sub1 <- subset.data.frame(LPC, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub2 <- subset.data.frame(LPS, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub3 <- subset.data.frame(LPF247, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub4 <- subset.data.frame(LPF112, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub5 <- subset.data.frame(LPF125, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub6 <- subset.data.frame(LPF242, select = c( "assay", "treatment", "date", "day", "plant_region_area"))

sub7 <- subset.data.frame(HPC, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub8 <- subset.data.frame(HPS, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub9 <- subset.data.frame(HPF247, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub10 <- subset.data.frame(HPF112, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub11 <- subset.data.frame(HPF125, select = c( "assay", "treatment", "date", "day", "plant_region_area"))
sub12 <- subset.data.frame(HPF242, select = c( "assay", "treatment", "date", "day", "plant_region_area"))



data_1 <- rbind(sub1, sub2)
data_2 <- rbind(data_1, sub3)
data_3 <- rbind(data_2, sub4)
data_4 <- rbind(data_3, sub5)
data_lowP <- rbind(data_4, sub6)

data_6 <- rbind(sub7, sub8)
data_7 <- rbind(data_6, sub9)
data_8 <- rbind(data_7, sub10)
data_9 <- rbind(data_8, sub11)
data_highP <- rbind(data_9, sub12)



rm(LPC)
rm(LPS)
rm(LPF247)
rm(LPF112)
rm(LPF125)
rm(LPF242)
rm(HPC)
rm(HPS)
rm(HPF247)
rm(HPF112)
rm(HPF125)
rm(HPF242)

rm(data_1)
rm(data_2)
rm(data_3)
rm(data_4)
rm(data_6)
rm(data_7)
rm(data_8)
rm(data_9)


```

```{r means, echo=F, message=F, warning=F }
#calculating mean values
means_sub1<- tapply(X=sub1$plant_region_area, INDEX=sub1$treatment, FUN=mean,na.rm=TRUE)#LPC
means_sub2<- tapply(X=sub2$plant_region_area, INDEX=sub2$treatment, FUN=mean,na.rm=TRUE)#LPS
means_sub3<- tapply(X=sub3$plant_region_area, INDEX=sub3$treatment, FUN=mean,na.rm=TRUE) #LPF247
means_sub4<- tapply(X=sub4$plant_region_area, INDEX=sub4$treatment, FUN=mean,na.rm=TRUE) #LPF112
means_sub5<- tapply(X=sub5$plant_region_area, INDEX=sub5$treatment, FUN=mean,na.rm=TRUE) #LPF125
means_sub6<- tapply(X=sub6$plant_region_area, INDEX=sub6$treatment, FUN=mean,na.rm=TRUE) #LPF242


rm(sub1)
rm(sub2)
rm(sub3)
rm(sub4)
rm(sub5)
rm(sub6)
rm(sub7)
rm(sub8)
rm(sub9)
rm(sub10)
rm(sub11)
rm(sub12)
```

ANOVA + Tukey's HSD
```{r anova last timepoint, echo=F, message=F, warning=F }
lowP<-aov(plant_region_area~treatment, data=data_lowP )
summary(lowP)

tukey_lowP<- TukeyHSD(lowP)
tukey_lowP


```
combine ANOVA + Tukey's HSD with compact letters
```{r,compact letter,echo=F, message=F, warning=F}

#plant_region_area
anova_plant_region_area<- aov(plant_region_area~treatment, data=data_lowP)
tukey_plant_region_area<-TukeyHSD(anova_plant_region_area)
cld1<- multcompLetters4(anova_plant_region_area, tukey_plant_region_area)
print(cld1)

Tk1<-group_by(data_lowP, treatment) %>%
  summarise(mean=mean(plant_region_area,na.rm=TRUE), quant=quantile(plant_region_area, probs=0.75, na.rm=TRUE)) %>%
  arrange(desc(mean))

cld1<- as.data.frame.list(cld1$treatment)
Tk1$cld1<-cld1$Letters
print(Tk1)

#plant_region_area
anova_plant_region_area<- aov(plant_region_area~treatment, data=data_highP)
tukey_plant_region_area<-TukeyHSD(anova_plant_region_area)
cld2<- multcompLetters4(anova_plant_region_area, tukey_plant_region_area)
print(cld2)

Tk2<-group_by(data_highP, treatment) %>%
  summarise(mean=mean(plant_region_area,na.rm=TRUE), quant=quantile(plant_region_area, probs=0.75, na.rm=TRUE)) %>%
  arrange(desc(mean))

cld2<- as.data.frame.list(cld1$treatment)
Tk2$cld2<-cld1$Letters
print(Tk2)


```

plotting
```{r,plot1,echo=F, message=F, warning=F}
#plant_region_area 
fig1<-data_lowP %>%
  ggplot(aes(x=treatment, y=plant_region_area, fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill=FALSE)+
  geom_text(data=Tk1, aes(label=cld1,x=treatment, y=quant),vjust=-5.7,hjust=-0.5, size=4)+
  scale_fill_manual(values = c("#cdc5c2", "#6699cc","#1974d2" ,"#17806d","#1a4876","#B2B2B2"))+
  scale_x_discrete("Treatment", labels=c("LP C", " LP F112", "LP F125", "LP F242","LP F247","LP Syn9"))+
    labs(x= "Treatment",
       y= "Rosette area (z-transformed)",
       title= "Rosette area under low P conditions")+
  scale_y_continuous(limits=c(0,4))+
  theme(legend.position = "none")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))
fig1
ggsave("boxplot_plant_region_area_LOWP.png", width=6, height=5) 

```


```{r,plot2,echo=F, message=F, warning=F}
#plant_region_area 
fig2<-data_highP %>%
  ggplot(aes(x=treatment, y=plant_region_area, fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill=FALSE)+
  geom_text(data=Tk2, aes(label=cld2,x=treatment, y=quant),vjust=-4.2,hjust=-0.5, size=4)+
  scale_fill_manual(values = c("#cdc5c2", "#6699cc","#1974d2" ,"#17806d","#1a4876","#B2B2B2"))+
  scale_x_discrete("Treatment", labels=c("HP C", " HP F112", "HP F125", "HP F242","HP F247","HP Syn9"))+
    labs(x= "Treatment",
       y= "Rosette area (z-transformed)",
       title= "Rosette area under high P conditions")+
  scale_y_continuous(limits=c(0,5))+
  theme(legend.position = "none")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))
fig2
ggsave("boxplot_plant_region_area_HIGHP.png", width=6, height=5) 

```