---
title: "Biomass Analysis: low and high P treatment"
author: "Claudia Probst"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    df_print: paged
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
--- 
# Experimental Setup

We're investigating the responses of *Arabidopsis thaliana* to varying levels of P-fertilizer supply in presence or absence of a synthetic microbial community. For a biomass analysis, plants are harvested at day 21 post germination. The plants are cut at the basis of the stem and the biomass from the shoot and root systems are quantified. For a dry weight analysis, the shoots and roots were dried at 60°C for 1-2 days. The experiment was conducted three times with the same growth conditions and the same phosphate treatment. 

\vspace{5mm}

Treatments: 

- low P control
- low P SynCom9
- high P control
- high P SynCom9
\vspace{5mm}

```{r setup, include=FALSE, echo=F, message=F, warning=F}

rm(list=ls())

library(ggplot2)
library(cowplot)
library(tibble)
library(tidyr)
library(ggsignif)
library(dplyr)
library(car)
library(RColorBrewer)
library(multcompView)
library(datasets)

setwd("~/Desktop/MASTER Pflanzenwissenschaften/Experiments/Magenta assay/MAGENTA LOW:HIGH")
data<-read.csv(file="MAGENTA_weight_low:highP.csv", header=TRUE, sep=",")

```

```{r mean, include=FALSE, echo=F, message=F, warning=F}
#calculation of mean 
means_shoot_fresh<- tapply(X=data$shoot_fresh_weight, INDEX=data$treatment, FUN=mean,na.rm=TRUE)
means_shoot_fresh

means_shoot_dry<- tapply(X=data$shoot_dry_weight, INDEX=data$treatment, FUN=mean,na.rm=TRUE)
means_shoot_dry

means_roots_fresh<- tapply(X=data$roots_fresh_weight, INDEX=data$treatment,FUN=mean,na.rm=TRUE)
means_roots_fresh

mean_roots_dry<-tapply(X=data$roots_dry_weight, INDEX = data$treatment, FUN=mean,na.rm=TRUE)
mean_roots_dry
```

```{r z_scores, echo=F, message=F, warning=F}
#features of interest
foi<- c("shoot_fresh_weight", "shoot_dry_weight", "roots_fresh_weight", "roots_dry_weight")
data_foi<- subset.data.frame(data, select=c("Assays", "treatment", "SynCom9", foi))
#prepare df
data_norm <- data_foi
data_norm[,(ncol(data_norm)-length(foi)+1):ncol(data_norm)] <- NA

#normalise by z-scores each feature for each time point
experiments <- unique(data_foi$Assays)

#loop
for (e in experiments) {
  for (f in foi) {
    x <- data_foi[data_foi$Assays == e, colnames(data_foi) == f] #get values for experiment 
    z <- (x-mean(x))/(sd(x)) #compute z_score
    data_norm[data_norm$Assays == e, colnames(data_norm) == f] <- z #store it
  }
}
```
# Statistical Analysis

### Normal distribution \| Shoot and Root biomass

Checking for normal distribution of the residuals
\vspace{5mm}

```{r normal distribuion, echo=F, message=F, warning=F, fig.width=10, fig.height=10}
# subsetting different treatment groups

#constant <- abs(min(data_norm$roots_dry_weight, na.rm = T))+1
#data_norm$roots_dry_weight_transformed <- #log(data_norm$roots_dry_weight + constant)

LPC<- subset(data_norm, treatment=="lowP_control")
HPC<- subset(data_norm, treatment=="highP_control")
LPS<- subset(data_norm, treatment=="lowP_SynCom9")
HPS<- subset(data_norm, treatment=="highP_SynCom9")


#low P control
par(mfrow=c(2,2))
for (i in 4:ncol(LPC)) {
  hist(LPC[[i]], main =colnames(LPC[i]))
}

par(mfrow=c(2,2))
for (i in 4:ncol(LPC)) {
  qqnorm((LPC[[i]]),main =colnames(LPC[i]))
  qqline((LPC[[i]]),main =colnames(LPC[i]))
}

shapiro.test(LPC$shoot_fresh_weight)
shapiro.test(LPC$shoot_dry_weight)
shapiro.test(LPC$roots_fresh_weight)
shapiro.test(LPC$roots_dry_weight)

#high P control
par(mfrow=c(2,2))
for (i in 4:ncol(HPC)) {
  hist(HPC[[i]], main =colnames(HPC[i]))
}

par(mfrow=c(2,2))
for (i in 4:ncol(HPC)) {
  qqnorm((HPC[[i]]),main =colnames(HPC[i]))
  qqline((HPC[[i]]),main =colnames(HPC[i]))
}

shapiro.test(HPC$shoot_fresh_weight)
shapiro.test(HPC$shoot_dry_weight)
shapiro.test(HPC$roots_fresh_weight)
shapiro.test(HPC$roots_dry_weight)

#low P SynCom9
par(mfrow=c(2,2))
for (i in 4:ncol(LPS)) {
  hist(LPS[[i]], main =colnames(LPS[i]))
}

par(mfrow=c(2,2))
for (i in 4:ncol(LPS)) {
  qqnorm((LPS[[i]]),main =colnames(LPS[i]))
  qqline((LPS[[i]]),main =colnames(LPS[i]))
}

shapiro.test(LPS$shoot_fresh_weight)
shapiro.test(LPS$shoot_dry_weight)
shapiro.test(LPS$roots_fresh_weight)
shapiro.test(LPS$roots_dry_weight)

#high P SynCom9
par(mfrow=c(2,2))
for (i in 4:ncol(HPS)) {
  hist(HPS[[i]], main =colnames(HPS[i]))
}

par(mfrow=c(2,2))
for (i in 4:ncol(HPS)) {
  qqnorm((HPS[[i]]),main =colnames(HPS[i]))
  qqline((HPS[[i]]),main =colnames(HPS[i]))
}

shapiro.test(HPS$shoot_fresh_weight)
shapiro.test(HPS$shoot_dry_weight)
shapiro.test(HPS$roots_fresh_weight)
shapiro.test(HPS$roots_dry_weight)

```

**Conclusion:** Samples seem to be normal distributed
\vspace{5mm}

### Anova \| Shoot and root biomass
```{r,anova,echo=F, message=F, warning=F}
A1<- aov(shoot_fresh_weight~treatment*Assays*SynCom9, data=data_norm)
summary(A1)
A2<- aov(shoot_dry_weight~treatment*Assays*SynCom9, data=data_norm)
summary(A2)
A3<- aov(roots_fresh_weight~treatment*Assays*SynCom9, data=data_norm)
summary(A3)
A4<- aov(roots_dry_weight~treatment*Assays*SynCom9, data=data_norm)
summary(A4)

anova1<- aov(shoot_fresh_weight~treatment, data=data_norm)
anova1
summary(anova1)
anova2<- aov(shoot_dry_weight~treatment, data=data_norm)
anova2
summary(anova2)
anova3<-aov(roots_fresh_weight~treatment, data=data_norm)
anova3
summary(anova3)
anova4<-aov(roots_dry_weight~treatment, data=data_norm)
anova4
summary(anova4)

```


```{r,Post-hoc-test,echo=F, message=F, warning=F, results=F}
tukey1<- TukeyHSD(anova1)
tukey1
tukey2<-TukeyHSD(anova2)
tukey2
tukey3<-TukeyHSD(anova3)
tukey3
tukey4<-TukeyHSD(anova4)
tukey4

```

```{r, compact letter display shoot fresh weight,echo=F, message=F, warning=F, results=F}
#shoot fresh weight
cld1<- multcompLetters4(anova1, tukey1)
print(cld1)

Tk1<- group_by(data_norm,treatment) %>%
  summarise(mean=mean(shoot_fresh_weight,na.rm=TRUE), quant=quantile(shoot_fresh_weight,na.rm=TRUE, probs=0.75))%>%
  arrange(desc(mean))

cld1<- as.data.frame.list(cld1$treatment)
Tk1$cld1<-cld1$Letters
print(Tk1)
```
#Figures 

### Figure 1 \| Shoot fresh weight
\vspace{5mm}
```{r, ggplot_shoot_freish_weight,echo=F, message=F, warning=F}

  
F1<-data_norm %>%
  ggplot(aes(x=treatment, y=shoot_fresh_weight, fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill=FALSE)+
  geom_text(data=Tk1, aes(label=cld1,x=treatment, y=quant),vjust=-4.6,hjust=-0.5, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"))+
  scale_shape_manual(values = c(16,17,15),labels=c("A1","A2","A3"))+
  labs( y= "shoot fresh weight (z-transformed)")+
  scale_y_continuous(limits = c(-2, 3))+
  scale_x_discrete(labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
  theme(legend.position = "bottom")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=9))+
  theme(text=element_text(size=12, colour="black"))+
  theme(plot.title = element_text(hjust=0.5))
F1
 # theme(legend.position = "TRUE")
#ggsave("boxplot_shoot_fresh_weight.png", width=6, height=4) 
```
\vspace{5mm}

**Conclusion:** Since sample size is different when comparing control treatments with SynCom9 treatments, we have to be careful with the interpretation. Also there is a lot of variation in the high P treatment. We can see, that there is a difference in the biomass when comparing control vs. SynCom9 treatment. The effect seem to be bigger when *Arabidopsis thaliana* is grown under P-stress. Different letters indicate a significant differences among groups (Post-hoc Tukey).

\vspace{5mm}

```{r, compact letter display shoot dry weight,echo=F, message=F, warning=F, results=F}

#shoot dry weight
cld2<- multcompLetters4(anova2, tukey2)
print(cld2)

Tk2<- group_by(data_norm,treatment) %>%
  summarise(mean=mean(shoot_dry_weight,na.rm=TRUE), quant=quantile(shoot_dry_weight,na.rm=TRUE, probs=0.75))%>%
  arrange(desc(mean))

cld2<- as.data.frame.list(cld2$treatment)
Tk2$cld2<-cld2$Letters
print(Tk2)
```

### Figure 2 \| Shoot dry weight
\vspace{5mm}
```{r message=FALSE, warning=FALSE, ggplot_shoot_dry_weight,echo=F}
#shoot dry weight
  
F2<-data_norm %>%
  ggplot(aes(x=treatment, y=shoot_dry_weight,fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill=FALSE)+
  geom_text(data=Tk2, aes(label=cld2,x=treatment, y=quant),vjust=-4,hjust=-0.5, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue","goldenrod1", "coral"))+
  scale_shape_manual(values = c(16,17,15),labels=c("A1","A2","A3"))+
  labs( y= "shoot dry weight (z-transformed)")+
  scale_y_continuous(limits = c(-2, 3))+
  scale_x_discrete(labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
  theme(legend.position = "bottom")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=9))+
  theme(text=element_text(size=12, colour="black"))+
  theme(plot.title = element_text(hjust=0.5))
F2
#ggsave("boxplot_shoot_dry_weight.png",width=6, height=4) 
```
\vspace{5mm}

**Conclusion:** Since sample size is different when comparing control treatments with SynCom9 treatments, we have to be careful with the interpretation. Also there is a lot of variation in the high P treatment. As above we can see a similar effect when comparing the treatments.

\vspace{5mm}

```{r, compact letter display roots fresh weight,echo=F, message=F, warning=F, results=F}

#roots fresh weight
cld3<- multcompLetters4(anova3, tukey3)
print(cld3)

Tk3<- group_by(data_norm,treatment) %>%
  summarise(mean=mean(roots_fresh_weight, na.rm=TRUE), quant=quantile(roots_fresh_weight, probs=0.75, na.rm=TRUE))%>%
  arrange(desc(mean))

cld3<- as.data.frame.list(cld3$treatment)
Tk3$cld3<-cld3$Letters
print(Tk3)
```
### Figure 3 \| Root fresh weight
\vspace{5mm}
```{r, ggplot_root_fresh_weight,echo=F, message=F, warning=F}

#root fresh weight

F3<-data_norm %>%
  ggplot(aes(x=treatment,y=roots_fresh_weight,fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill=FALSE)+
  geom_text(data=Tk3, aes(label=cld3,x=treatment, y=quant),vjust=-3.5,hjust=-0.8, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue","goldenrod1", "coral"))+
  scale_shape_manual(values = c(16,17,15),labels=c("A1","A2","A3"))+
  labs( y= "root fresh weight (z-transformed)")+
  scale_y_continuous(limits = c(-2, 3))+
  scale_x_discrete(labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
  theme(legend.position = "bottom")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=9))+
  theme(text=element_text(size=12, colour="black"))+
  theme(plot.title = element_text(hjust=0.5))
F3
#ggsave("boxplot_roots_fresh_weight.png",width=6, height=4) 
```

\vspace{5mm}

**Conclusion:** Since sample size is different when comparing control treatments with SynCom9 treatments, we have to be careful with the interpretation. Also there is a lot of variation in the all treatments. We can see a trend, that *Arabidopsis thaliana* grown under low P condition and with microbial treatments, have a higher biomass compared to sterile low P conditions.  

\vspace{5mm}
```{r, compact letter display roots dry weight,echo=F, message=F, warning=F, results=F}

#roots dry weight
cld4<- multcompLetters4(anova4, tukey4)
print(cld4)

Tk4<- group_by(data_norm,treatment) %>%
  summarise(mean=mean(roots_dry_weight, na.rm=TRUE), quant=quantile(roots_dry_weight, probs=0.75, na.rm=TRUE))%>%
  arrange(desc(mean))

cld4<- as.data.frame.list(cld4$treatment)
Tk4$cld4<-cld4$Letters
print(Tk4)

```
### Figure 4 \| Root dry weight
\vspace{5mm}
```{r, ggplot_root_dry_weight,echo=F, message=F, warning=F}
#root dry weight

F4<-data_norm %>%
  ggplot(aes(x=treatment,y=roots_dry_weight,fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+ 
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill=FALSE)+
  geom_text(data=Tk4, aes(label=cld4,x=treatment, y=quant),vjust=-5.2,hjust=-0.5, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue","goldenrod1", "coral"))+
  scale_shape_manual(values = c(16,17,15),labels=c("A1","A2","A3"))+
  labs( y= "root dry weight (z-transformed)")+
  scale_y_continuous(limits = c(-2, 3))+
  scale_x_discrete(labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
  theme(legend.position = "bottom")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=9))+
  theme(text=element_text(size=12, colour="black"))+
  theme(plot.title = element_text(hjust=0.5))
F4
#ggsave("boxplot_roots_dry_weight.png",width=6, height=4) 
```
\vspace{5mm}

**Conclusion:** Since sample size is different when comparing control treatments with SynCom9 treatments, we have to be careful with the interpretation. Also there is a lot of variation in the all treatments. 

\vspace{5mm}

```{r, plot_grid,echo=F, message=F, warning=F}
fig <- plot_grid(F1 + labs(colour=""),
                F2 + labs(colour=""),
                F3 + labs(colour=""),
                F4 + labs(colour=""),
               labels=c("A","B","C","D"),
               align = "hv",
               rel_heights = c(1,1),
               nrow=2, ncol=2, label_size = 11)

legend<- get_legend(F1 +theme(legend.position = "bottom")+ labs(fill="treatment")+theme(legend.title = ))

title<- ggdraw()+draw_label("Biomass effects by SynCom9", fontface = "bold")


FIG<- plot_grid(fig,legend, nrow=2,ncol=1,rel_heights=c( 1,.1 ))

#ggsave("FIG1.png",width=6, height=6) 


```



