---
title: "NBRIP SI analysis"
author: "Claudia Probst"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
---
\pagebreak

# Experimental Setup

Different phosphate sources were used to compare the capacity of the individual strains to solubilize inorganic phosphate. National Botanical Research Institute’s phosphate growth medium (NBRIP) supplemented with bromophenol  blue (NBRIP-BPB) served as basis medium (5 g L-1 MgCl2 x 6H2O, 0.25 g L-1 MgSO4 x H2O, 0.2 g L-1 KCl, 0.1 g L-1 (NH4)2SO4, 10 g L-1 D(+)-glucose , ± 0.025 g L-1 bromophenol blue (BPB), 12 g L-1 agar, pH 7.0) (Nautiyal, 1999, Gupta et al., 1994). Beside tricalcium phosphate (5 g L-1 Ca3(PO4)2), dipotassium phosphate (K2HPO4) with 1mM or 10mM concentrations served as phosphate sources. Furthermore, we investigated the growth of bacteria in absence of phosphorous (no P). 5 x 10 µl of inoculum were spotted on each plate and incubated in the dark for 2-9 days. The quantification was done with the solubilization index (SI) according to Edi-Premono (Edi-Premono et al., 1996): SI=(colony diameter+halo zone diameter)/(colony diameter). Strains forming a clear halo zone around the colony were identified as P-solubilizer. The experiment was conducted in triplicates. 

\vspace{5mm}

Treatments:

- TSB: control for bacterial growth
- Ca3(PO4)2
- 1mM K2HPO4: control for halo formation
- 10mM K2HPO4
- no phosphate: control for halo formation
\vspace{5mm}

```{r setup, include=FALSE, echo=F, message=F, warning=F}


library(ggplot2)
library(cowplot)
library(tibble)
library(tidyr)
library(ggsignif)
library(dplyr)
library(car)
library(RColorBrewer)
library(multcompView)
library(datasets)
library(grid)
library(gridExtra)
library(broman)
library(dunn.test)

 
##clear the objects from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)

#set seed
set.seed(100)

# installs (if necessary) and loads libraries
library("readxl")

#import
setwd("~/Desktop/MASTER Pflanzenwissenschaften/Experiments/NBRIP assay")
NBRIP_Ca<- read_excel("NBRIP_data_all.xlsx", sheet = 1)
NBRIP_10mM<- read_excel("NBRIP_data_all.xlsx", sheet = 2)


```


```{r foi, echo=F, message=F, warning=F}

# select features of interest
ca<- subset.data.frame(NBRIP_Ca, select = c( "strain","SI_diameter", "SI_area", "assay"))
k2<- subset.data.frame(NBRIP_10mM, select = c( "strain","SI_diameter", "SI_area","assay"))

#ca_solubilizer<- subset(ca, strain %in% c("F247", "F112","F242","control"))
#k2_solubilizer<-subset(k2, strain %in% c("F247", "F112", "F242", "F201","F004","control"))
```


```{r mean, include=FALSE, echo=F, message=F, warning=F}
#calculation of mean Ca3(PO4)2
ca_means_SI_diameter<- tapply(X=ca$SI_diameter, INDEX=ca$strain, FUN=mean, na.rm=TRUE)
ca_means_SI_diameter

#calculation of mean 10mM K2HPO4

k2_means_SI_diameter<- tapply(X=k2$SI_diameter, INDEX=k2$strain, FUN=mean, na.rm=TRUE)
k2_means_SI_diameter

```

```{r PSB, include=FALSE, echo=F, message=F, warning=F}

#create a new variable PSB Ca3(PO4)2

ca$PSB<-factor(NA, levels =c("effective PSB","weak PSB", "no PSB"))
ca$PSB[ca$SI_diameter >= 1.5] <- "effective PSB"
ca$PSB[ca$SI_diameter> 1 & ca$SI_diameter < 1.5] <- "weak PSB"
ca$PSB[ca$SI_diameter ==1] <- "no PSB"

#create a new variable PSB K2HPO4

k2$PSB<-factor(NA, levels =c("effective PSB","weak PSB", "no PSB"))
k2$PSB[k2$SI_diameter >= 1.5] <- "effective PSB"
k2$PSB[k2$SI_diameter> 1 & k2$SI_diameter < 1.5] <- "weak PSB"
k2$PSB[k2$SI_diameter ==1] <- "no PSB"
```

```{r dataframe, include=FALSE, echo=F, message=F, warning=F}

#create new dataframe with mean values and PSB Ca3(PO4)2

df_ca<- data.frame(ca_means_SI_diameter)
df_ca$PSB<-factor(NA, levels =c("effective PSB","weak PSB", "no PSB"))
df_ca$PSB[df_ca$ca_means_SI_diameter >= 1.5] <- "effective PSB"
df_ca$PSB[df_ca$ca_means_SI_diameter> 1 & df_ca$ca_means_SI_diameter < 1.5] <- "weak PSB"
df_ca$PSB[df_ca$ca_means_SI_diameter ==1] <- "no PSB"
df_ca

#create new dataframe with mean values and PSB 10mM K2HPO4

df_k2<- data.frame(k2_means_SI_diameter)
df_k2$PSB<-factor(NA, levels =c("effective PSB","weak PSB", "no PSB"))
df_k2$PSB[df_k2$k2_means_SI_diameter >= 1.5] <- "effective PSB"
df_k2$PSB[df_k2$k2_means_SI_diameter> 1 & df_k2$k2_means_SI_diameter < 1.5] <- "weak PSB"
df_k2$PSB[df_k2$k2_means_SI_diameter ==1] <- "no PSB"
df_k2
```

# Statistical Analysis

### Normal distribtion \| NBRIP Ca3(PO4)2

Checking for normal distribution of the residuals

* SI diameter
* SI area

\vspace{5mm}

```{r CA normal distribuion +Anova, echo=F, message=F, warning=F}
#SI diameter normal distribution
par(mfrow=c(1:2))
hist(ca$SI_diameter, xlab="SI diameter", ylab="frequency", main="Histogram")
hist(rank(ca$SI_diameter), xlab="SI diameter", ylab="frequency", main="Histogram")
qqplot<-scale(ca$SI_diameter)
qqnorm(qqplot)
qqline(qqplot)


#SI area normal distribution
par(mfrow=c(1:2))
hist(ca$SI_area, xlab="SI area", ylab="frequency", main="Histogram")
hist(rank(ca$SI_area), xlab="SI area", ylab="frequency", main="Histogram")
qqplot<-scale(ca$SI_area)
qqnorm(qqplot)
qqline(qqplot)

```
\vspace{5mm}
**Conclusion:** Samples are not normal distributed. Therfore we apply a Kruksal Wallis test. 
\vspace{5mm}

### Kruksal Wallis test \| NBRIP Ca3(PO4)2
```{r,kruksal wallis CA,echo=F, message=F, warning=F}

KW1<-as.factor(paste(ca$strain, ca$assay,sep="_"))
sample<-ca$SI_diameter
Kruk1<-kruskal.test(sample~KW1)
Kruk1


KW2<-as.factor(paste(k2$strain, k2$assay,sep="_"))
sample<-k2$SI_diameter
Kruk2<-kruskal.test(sample~KW2)
Kruk2

```
\vspace{5mm}
**Conclusion:** In both media we see a significant difference between the strains. 
\vspace{5mm}

### Dunn-Sidak test \| NBRIP Ca3(PO4)2
```{r,dunn.test CA,echo=F, message=F, warning=F, results=F, list=TRUE}
library(dunn.test)
library(rstatix)

dt1<-dunn.test(ca$SI_diameter,ca$strain,method="bonferroni", kw=TRUE, wrap=TRUE)
stat.test1<-dunn_test(SI_diameter~strain, data=ca)

dt2<-dunn.test(ca$SI_area, ca$strain, method="bonferroni", kw=TRUE, wrap=TRUE)
stat.test2<-dunn_test(SI_area~strain, data=ca)

dt3<-dunn.test(k2$SI_diameter, k2$strain, method="bonferroni", kw=TRUE, wrap=TRUE)
stat.test3<-dunn_test(SI_diameter~strain, data=k2)

dt4<-dunn.test(k2$SI_area, k2$strain, method="bonferroni", kw=TRUE, wrap=TRUE)
stat.test4<-dunn_test(SI_area~strain, data=k2)


```
\vspace{5mm}
**Conclusion:** By comparing all the groups against each other, groups which differ from each other are considered as P-solubilizer. 
\vspace{5mm}

### Figure 1\| Ca3(PO4)2 solubilization by SynCom9 members
\vspace{5mm}
```{r, ggplot1,echo=F, message=F, warning=F}
library(ggpubr)
library(multcompView)
library(rcompanion)

letters <- cldList(P.adjusted ~ comparisons, data=dt1, threshold=0.05)
letters$Group <- gsub("^F4$", "F004", letters$Group)
letters$Group <- gsub("^F41$", "F041", letters$Group)
letters$Group <- gsub("^F42$", "F042", letters$Group)
letters$Group <- gsub("^F21$", "F201", letters$Group)

ca$letter <- NA
for (s in unique(ca$strain)) {ca$letter[ca$strain==s] <- letters$Letter[letters$Group==s]}


plot1<- ca %>%
  ggplot(aes(x=strain, y=SI_diameter, fill=strain))+
  #geom_text(aes(label=letter, y=0.8),check_overlap=TRUE)+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.6)+
  geom_jitter(aes(shape=assay), width = 0.3, alpha=0.7)+
  theme_classic()+
  guides(fill=FALSE)+
  scale_fill_manual(values = c("#cdc5c2","#BF84D3","#fce883","#ffcf48", "#6699cc","#1974d2" ,"#324ab2", "#3bb08f","#17806d","#1a4876" ))+
  labs(x= "Strains",
       y= "SI diameter")+
  theme(text = element_text(size=12))+
  theme(axis.text.x = element_text(size = 11))+
  theme(axis.text.y = element_text(size = 11))+
  theme(legend.position = "none")+
  #stat_compare_means(method="kruskal.test", label.x=1, label.y=4)+
  annotate("text", x=1, y=1.2, label="a")+
  annotate("text", x=2, y=1.2, label="a")+
  annotate("text", x=3, y=1.2, label="a")+
  annotate("text", x=4, y=1.2, label="a")+
  annotate("text", x=5.2, y=2.2, label="b")+
  annotate("text", x=6, y=1.2, label="a")+
  annotate("text", x=7, y=1.2, label="a")+
  annotate("text", x=8, y=1.2, label="a")+
  annotate("text", x=9.2, y=2.8, label="b")+
  annotate("text", x=10.2, y=2.15, label="b")+
  ggtitle("Solubilization of tricalcium phosphate by SynCom9 members")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=10))+
  theme(text=element_text(size=12, colour="black"))


```

```{r, ggplot2,echo=F, message=F, warning=F, fig.height=8, fig.width=10}


plot2<-ggplot(ca,aes(x=strain, y=SI_area, fill=strain))+
 # geom_text(aes(label=letter, y=0.6),check_overlap=TRUE)+
  geom_boxplot(outlier.shape = NA,alpha=0.6,width = 0.6)+
  geom_jitter(aes(shape=assay), width = 0.3, alpha=0.7)+
  theme_classic()+
  guides(fill=FALSE)+
  scale_fill_manual(values = c("#cdc5c2","#BF84D3","#fce883","#ffcf48", "#6699cc","#1974d2" ,"#324ab2", "#3bb08f","#17806d","#1a4876" ))+
  labs(x= "Strains",
       y= "SI area")+
  theme(text = element_text(size=12))+
  theme(axis.text.x = element_text(size =11))+
  theme(axis.text.y = element_text(size =11))+
  theme(legend.position = "bottom")+
  stat_compare_means(method="kruskal.test", label.x=1, label.y=16)+
  annotate("text", x=5, y=5, label="****")+
  annotate("text", x=9, y=16, label="****")+
  annotate("text", x=10, y=16, label="****")+
  ggtitle("Solubilization of tricalcium phosphate by SynCom9 members")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=10))+
  theme(text=element_text(size=12, colour="black"))


fig1 <- plot_grid(plot1, plot2, labels=c("A","B"), ncol=1)
title<- ggdraw()+draw_label("Solubilization of Tricalcium phosphate by SynCom9 members", fontface = "bold")
plot_grid(title,fig1, ncol=1,rel_heights = c(0.1,1))

ggsave("fig1.png", height=8, width = 10)


```

\vspace{5mm}
**Conclusion:** F112 (Pseudomonas), F242 (Labrys), F247 (Pseudomonas) were able to form a halo around the colony and thus identified as P-solubilizer. The higher the SI value, the more calcium phosphate was solubilized. Values around 1 indicate strains that were not able to solubilize phosphorous.Different symbols indicate replicates.
\vspace{5mm}


### Normal distribtion \| NBRIP 10mM K2HPO4

Checking for normal distribution of the residuals
*SI diameter*
*SI area*

\vspace{5mm}

```{r K2 normal distribuion , echo=F, message=F, warning=F}
#SI diameter normal distribution
par(mfrow=c(1:2))
diameter<- aov(k2$SI_diameter~k2$strain*k2$assay)
hist(rstandard(diameter))
plot(diameter,2)

#SI area normal distribution
par(mfrow=c(1:2))
area<- aov(k2$SI_area~k2$strain*k2$assay)
hist(rstandard(area))
plot(area,2)


```

\vspace{5mm}
**Conclusion:** Samples are not normal distributed.
\vspace{5mm}

\vspace{5mm}
### Figure 2\| K2HPO4 solubilization by SynCom9 members
```{r, ggplot3,echo=F, message=F, warning=F}
letters <- cldList(P.adjusted ~ comparisons, data=dt3, threshold=0.05)
letters$Group <- gsub("^F4$", "F004", letters$Group)
letters$Group <- gsub("^F41$", "F041", letters$Group)
letters$Group <- gsub("^F42$", "F042", letters$Group)
letters$Group <- gsub("^F21$", "F201", letters$Group)

k2$letter <- NA
for (s in unique(k2$strain)) {k2$letter[k2$strain==s] <- letters$Letter[letters$Group==s]}


plot3<-ggplot(k2,aes(x=strain, y=SI_diameter, fill=strain))+
#  geom_text(aes(label=letter, y=0.8),check_overlap=TRUE)+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.6)+
  geom_jitter(aes(shape=assay), width = 0.3, alpha=0.7)+
  theme_classic()+
  guides(fill=FALSE)+
  scale_fill_manual(values = c("#cdc5c2","#BF84D3","#fce883","#ffcf48", "#6699cc","#1974d2" ,"#324ab2", "#3bb08f","#17806d","#1a4876" ))+
  labs(x= "Strains",
       y= "SI diameter")+
  theme(text = element_text(size=12))+
  theme(axis.text.x = element_text(size = 11))+
  theme(axis.text.y = element_text(size = 11))+
  theme(legend.position = "none")+
  #stat_compare_means(method="kruskal.test", label.x=1, label.y=5)+
  annotate("text", x=1, y=1.2, label="a")+
  annotate("text", x=2.2, y=2.1, label="b")+
  annotate("text", x=3, y=1.2, label="a")+
  annotate("text", x=4, y=1.2, label="a")+
  annotate("text", x=5.2, y=4.4, label="c")+
  annotate("text", x=6, y=1.2, label="a")+
  annotate("text", x=7.2, y=1.8, label="ab")+
  annotate("text", x=8, y=1.2, label="a")+
  annotate("text", x=9.2, y=3.6, label="bc")+
  annotate("text", x=10.2, y=3.8, label="bc")+
  ggtitle("Solubilization of dipotassium phoshate by SynCom9 members")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=10))+
  theme(text=element_text(size=12, colour="black"))
  

```

```{r, ggplot4,echo=F, message=F, warning=F, fig.height=8, fig.width=10}

plot4<-ggplot(k2,aes(x=strain, y=SI_area, fill=strain))+
 # geom_text(aes(label=letter, y=0.8),check_overlap=TRUE)+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.6)+
  geom_jitter(aes(shape=assay), width = 0.3, alpha=0.7)+
  theme_classic()+
  guides(fill=FALSE)+
  scale_fill_manual(values = c("#cdc5c2","#BF84D3","#fce883","#ffcf48", "#6699cc","#1974d2" ,"#324ab2", "#3bb08f","#17806d","#1a4876" ))+
  labs(x= "Strains",
       y= "SI diameter")+
  theme(text = element_text(size=12))+
  theme(axis.text.x = element_text(size = 11))+
  theme(axis.text.y = element_text(size = 11))+
  theme(legend.position = "bottom")+
  stat_compare_means(method="kruskal.test", label.x=1, label.y=19)+
  annotate("text", x=2, y=5, label="*")+
  annotate("text", x=5, y=18, label="****")+
  annotate("text", x=7, y=6, label="ns")+
  annotate("text", x=9, y=13, label="****")+
  annotate("text", x=10, y=14.5, label="****")+
  ggtitle("Solubilization of dipotassium phoshate by SynCom9 members")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=10))+
  theme(text=element_text(size=12, colour="black"))


legend<- get_legend(plot1 +theme(legend.position = "bottom")+ labs(fill="assay")+theme(legend.title = ))
fig <- plot_grid(plot1, plot3, labels=c("A","B"), ncol=1)
#title<- ggdraw()+draw_label("Solubilization of Dipotassium phosphate by SynCom9 members", fontface = "bold")
plot_grid(fig,legend, nrow=2,ncol=1,rel_heights=c( 1,.1 ))

ggsave("fig.png", height=7, width = 8)

```
\vspace{5mm}
**Conclusion:** F004 (Domibacillus), F112 (Pseudomonas), F201 (Ensifer), F242 (Labrys), F247 (Pseudomonas) were able to form a halo around the colony and thus identified as P-solubilizer. The higher the SI value, the more dipotassium phosphate was solubilized. Values around 1 indicate strains that were not able to solubilize phosphorous. Different symbols indicate replicates. 
\vspace{5mm}
