---
title: "Image Analysis low/highP"
author: "Claudia Probst"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
---
# Experimental Setup

We're investigating the responses of *Arabidopsis thaliana* to varying levels of P-fertilizer supply in presence or absence of a synthetic microbial community. ARADEEPOPSIS is a software that can be used to study pheotypic traits of *Arabidopsis thaliana* by images. We can investigate biomass, shape and size by different features over time. The experiment was conducted three times with the same growth conditions and the same phosphate treatments. 

```{r setup, include=FALSE, echo=F, message=F, warning=F}

##clear the objects from memory
rm(list=ls())


## installs (if necessary) and loads libraries
#install.packages("ggplot2")
#install.packages("ggthemes")
#install.packages("ggpubr")
#install.packages("ggsignif")
#install.packages("cowplot")
#install.packages("tibble")
#install.packages("tidyr")
#install.packages("tidyverse")
#install.packages("datasets")
#install.packages("multcomp")
#install.packages("multcompView")
#install.packages("dplyr")
#install.packages("plyr")   
#install.packages("readr")   

library(ggplot2)
library(ggthemes)
library(ggpubr)
library(cowplot)
library(tibble)
library(tidyr)
library(tidyverse)
library(ggsignif)
library(datasets)
library(multcomp)
library(multcompView)
library(dplyr)   
#library(plyr)  
library(readr) 

#set seed
set.seed(100)

#import
setwd("~/Desktop/MASTER Pflanzenwissenschaften/Experiments/Magenta assay/MAGENTA LOW:HIGH")

magenta2<-read.csv(file="magenta2.csv", header=TRUE, sep=",")
magenta3<-read.csv(file="magenta3.csv", header=TRUE, sep=",")
magenta4<-read.csv(file="magenta4.csv", header=TRUE, sep=",")
magenta5<-read.csv(file="magenta5.csv", header=TRUE, sep=",")

```

```{r combine datasets, include=FALSE, echo=F, message=F, warning=F}

#combine all data
data_1 <- rbind(magenta2, magenta3)
data_2 <- rbind(data_1, magenta4)
data_total <- rbind(data_2, magenta5)

str(data_total)

rm(magenta2)
rm(magenta3)
rm(magenta4)
rm(magenta5)
rm(data_1)
rm(data_2)
```

```{r foi, echo=F, message=F, warning=F}

# select features of interest
foi <- c("plant_region_red_channel","plant_region_green_channel", "plant_region_blue_channel", "plant_region_chroma_ratio", "plant_region_chroma_difference", "plant_region_chroma_base", "plant_region_green_strength","plant_region_blue_green_ratio", "plant_region_area", "plant_region_filled_area", "plant_region_convex_area","plant_region_equivalent_diameter", "plant_region_major_axis_length","plant_region_minor_axis_length", "plant_region_perimeter","plant_region_eccentricity", "plant_region_extent", "plant_region_solidity", "plant_region_aspect_ratio")
data_foi <- subset.data.frame(data_total, select = c("file", "Assays", "treatment", "date", "day", foi))

```

```{r z_scores, echo=F, message=F, warning=F}
#prepare df
data_foi_norm <- data_foi
data_foi_norm[,(ncol(data_foi_norm)-length(foi)+1):ncol(data_foi_norm)] <- NA

#normalise by z-scores each feature for each assay
assays <- unique(data_foi$Assays)

#loop
for (e in assays) {
  for (f in foi) {
    x <- data_foi[data_foi$Assays == e, colnames(data_foi) == f] #get values for experiment for foi
    z <- (x-mean(x))/(sd(x)) #compute z_score
    data_foi_norm[data_foi_norm$Assays == e, colnames(data_foi_norm) == f] <- z #store it
  }
}

rm(data_foi)
```
# Statistical Analysis

### Normal distribution \| Plant biomass analysis at day 21

Checking for normal distribution of the residuals

\vspace{5mm}

```{r final dataset,echo=F, message=F, warning=F}
all_data<- subset(data_foi_norm, day=="day 21")

```

```{r normal distribution last timepoint,echo=F, message=F, warning=F, fig.show="hide"}

par(mfrow=c(1:2))
for (i in 6:ncol(all_data)) {
  hist(all_data[[i]], main =colnames(all_data[i]))
}

LPC<- subset(all_data, treatment=="lowP_control")
HPC<- subset(all_data, treatment=="highP_control")
LPS<- subset(all_data, treatment=="lowP_SynCom9")
HPS<- subset(all_data, treatment=="highP_SynCom9")


#low P control
par(mfrow=c(2,2))
for (i in 6:ncol(LPC)) {
  hist(LPC[[i]], main =colnames(LPC[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(LPC)) {
  qqnorm((LPC[[i]]),main =colnames(LPC[i]))
  qqline((LPC[[i]]),main =colnames(LPC[i]))
}

shapiro.test(LPC$plant_region_area)
shapiro.test(LPC$plant_region_green_strength)
shapiro.test(LPC$plant_region_major_axis_length)

#log transformation
constant <- abs(min(LPC$plant_region_area, na.rm = T))+1
LPC$plant_region_area <- log(LPC$plant_region_area + constant)

constant <- abs(min(LPC$plant_region_green_strength, na.rm = T))+1
LPC$plant_region_green_strength <- log(LPC$plant_region_green_strength + constant)

constant <- abs(min(LPC$plant_region_major_axis_length, na.rm = T))+1
LPC$plant_region_major_axis_length <- log(LPC$plant_region_major_axis_length + constant)

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  hist(LPS[[i]], main =colnames(LPS[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  qqnorm((LPS[[i]]),main =colnames(LPS[i]))
  qqline((LPS[[i]]),main =colnames(LPS[i]))
}

shapiro.test(LPC$plant_region_area)
shapiro.test(LPC$plant_region_green_strength)
shapiro.test(LPC$plant_region_major_axis_length)

#high P control
par(mfrow=c(2,2))
for (i in 6:ncol(HPC)) {
  hist(HPC[[i]], main =colnames(HPC[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(HPC)) {
  qqnorm((HPC[[i]]),main =colnames(HPC[i]))
  qqline((HPC[[i]]),main =colnames(HPC[i]))
}

shapiro.test(HPC$plant_region_area)
shapiro.test(HPC$plant_region_green_strength)
shapiro.test(HPC$plant_region_major_axis_length)

#log transformation
constant <- abs(min(HPC$plant_region_area, na.rm = T))+1
HPC$plant_region_area <- log(HPC$plant_region_area + constant)

constant <- abs(min(HPC$plant_region_green_strength, na.rm = T))+1
HPC$plant_region_green_strength <- log(HPC$plant_region_green_strength + constant)

constant <- abs(min(HPC$plant_region_major_axis_length, na.rm = T))+1
HPC$plant_region_major_axis_length <- log(HPC$plant_region_major_axis_length + constant)

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  hist(LPS[[i]], main =colnames(LPS[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  qqnorm((LPS[[i]]),main =colnames(LPS[i]))
  qqline((LPS[[i]]),main =colnames(LPS[i]))
}

shapiro.test(HPC$plant_region_area)
shapiro.test(HPC$plant_region_green_strength)
shapiro.test(HPC$plant_region_major_axis_length)


#low P SynCom9
par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  hist(LPS[[i]], main =colnames(LPS[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  qqnorm((LPS[[i]]),main =colnames(LPS[i]))
  qqline((LPS[[i]]),main =colnames(LPS[i]))
}

shapiro.test(LPS$plant_region_area)
shapiro.test(LPS$plant_region_green_strength)
shapiro.test(LPS$plant_region_major_axis_length)

#log transformation
constant <- abs(min(LPS$plant_region_area, na.rm = T))+1
LPS$plant_region_area <- log(LPS$plant_region_area + constant)

constant <- abs(min(LPS$plant_region_green_strength, na.rm = T))+1
LPS$plant_region_green_strength <- log(LPS$plant_region_green_strength + constant)

constant <- abs(min(LPS$plant_region_major_axis_length, na.rm = T))+1
LPS$plant_region_major_axis_length <- log(LPS$plant_region_green_strength + constant)


par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  hist(LPS[[i]], main =colnames(LPS[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  qqnorm((LPS[[i]]),main =colnames(LPS[i]))
  qqline((LPS[[i]]),main =colnames(LPS[i]))
}

shapiro.test(LPS$plant_region_area)
shapiro.test(LPS$plant_region_green_strength)
shapiro.test(LPS$plant_region_major_axis_length)



#high P SynCom9
par(mfrow=c(2,2))
for (i in 6:ncol(HPS)) {
  hist(HPS[[i]], main =colnames(HPS[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(HPS)) {
  qqnorm((HPS[[i]]),main =colnames(HPS[i]))
  qqline((HPS[[i]]),main =colnames(HPS[i]))
}

shapiro.test(HPS$plant_region_area)
shapiro.test(HPS$plant_region_green_strength)
shapiro.test(HPS$plant_region_major_axis_length)

#log transformation
constant <- abs(min(HPS$plant_region_area, na.rm = T))+1
HPS$plant_region_area <- log(HPS$plant_region_area + constant)

constant <- abs(min(HPS$plant_region_green_strength, na.rm = T))+1
HPS$plant_region_green_strength <- log(HPS$plant_region_green_strength + constant)

constant <- abs(min(HPS$plant_region_major_axis_length, na.rm = T))+1
HPS$plant_region_major_axis_length <- log(HPS$plant_region_green_strength + constant)

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  hist(LPS[[i]], main =colnames(LPS[i]))
}

par(mfrow=c(2,2))
for (i in 6:ncol(LPS)) {
  qqnorm((LPS[[i]]),main =colnames(LPS[i]))
  qqline((LPS[[i]]),main =colnames(LPS[i]))
}

shapiro.test(HPS$plant_region_area)
shapiro.test(HPS$plant_region_green_strength)
shapiro.test(HPS$plant_region_major_axis_length)

```

```{r merging subsets, echo=F, message=F, warning=F }

#merging all subsets to one dataframe

sub1 <- subset.data.frame(HPS, select = c( "Assays", "treatment", "date", "day", "plant_region_area", "plant_region_major_axis_length", "plant_region_green_strength"))

sub2 <- subset.data.frame(HPC, select = c( "Assays", "treatment", "date", "day", "plant_region_area", "plant_region_major_axis_length", "plant_region_green_strength"))

sub3 <- subset.data.frame(LPS, select = c( "Assays", "treatment", "date", "day", "plant_region_area", "plant_region_major_axis_length", "plant_region_green_strength"))

sub4 <- subset.data.frame(LPC, select = c( "Assays", "treatment", "date", "day", "plant_region_area", "plant_region_major_axis_length", "plant_region_green_strength"))


data_1 <- rbind(sub1, sub2)
data_2 <- rbind(data_1, sub3)
data_total <- rbind(data_2, sub4)

rm(HPC)
rm(HPS)
rm(LPC)
rm(LPS)


#calculating mean values
means_sub1<- tapply(X=sub1$plant_region_area, INDEX=sub1$treatment, FUN=mean,na.rm=TRUE)
means_sub2<- tapply(X=sub2$plant_region_area, INDEX=sub2$treatment, FUN=mean,na.rm=TRUE)
means_sub3<- tapply(X=sub3$plant_region_area, INDEX=sub3$treatment, FUN=mean,na.rm=TRUE) #LPS
means_sub4<- tapply(X=sub4$plant_region_area, INDEX=sub4$treatment, FUN=mean,na.rm=TRUE) #LPC


# Calculate the percentage difference between the means of the two treatments
percentage_diff1 <- abs(means_sub1 - means_sub2) / mean(c(means_sub1, means_sub2)) * 100

percentage_diff2 <- abs(means_sub3 - means_sub4) / mean(c(means_sub3, means_sub4)) * 100

# Print the percentage difference
cat("The percentage difference between high P control and high P SynCom9 is", round(percentage_diff1, 2), "%")

cat("The percentage difference between low P control and low P SynCom9 is", round(percentage_diff2, 2), "%")

```

### Anova \| Size / Color / Shape
```{r,anova,echo=F, message=F, warning=F}
a1<- aov(plant_region_area~treatment*Assays, data=data_total)
summary(a1)
a2<- aov(plant_region_green_strength~treatment*Assays, data=data_total)
summary(a2)
a3 <- aov(plant_region_major_axis_length~treatment*Assays, data=data_total)
summary(a3)

anova1<- aov(plant_region_area~treatment, data=data_total)
anova1
summary(anova1)
anova2<- aov(plant_region_green_strength~treatment, data=data_total)
anova2
summary(anova2)
anova3<-aov(plant_region_major_axis_length~treatment, data=data_total)
anova3
summary(anova3)


```


```{r,Post-hoc-test,echo=F, message=F, warning=F, results=F}
tukey1<- TukeyHSD(anova1)
tukey1
tukey2<-TukeyHSD(anova2)
tukey2
tukey3<-TukeyHSD(anova3)
tukey3

```

GRAPHICAL ANALYIS: ggplot_boxplot with COMPACT LETTERS
ANOVA and post hoc Tukey test for 3 features of interest (all data --> data foi norm):
We will chose 3 features of interest (we have to calculate Anova and Tukey values again, to get compact letters): 
- for size: plant_region_area
- for shape: plant_region_major_axis_length
- for colour: plant_region_chroma_ratio

```{r,compact letter_pra,echo=F, message=F, warning=F, results=F}

#plant_region_area
anova_plant_region_area<- aov(plant_region_area~treatment, data=data_total)
tukey_plant_region_area<-TukeyHSD(anova_plant_region_area)
cld1<- multcompLetters4(anova_plant_region_area, tukey_plant_region_area)
print(cld1)

Tk1<-group_by(data_total, treatment) %>%
  summarise(mean=mean(plant_region_area,na.rm=TRUE), quant=quantile(plant_region_area, probs=0.75, na.rm=TRUE)) %>%
  arrange(desc(mean))

cld1<- as.data.frame.list(cld1$treatment)
Tk1$cld1<-cld1$Letters
print(Tk1)
```
# Graphical analysis

### Figure 1 \| Plant region area at day 21

\vspace{5mm}

```{r,plot_pra,echo=F, message=F, warning=F}

#plant_region_area
F1<-data_total %>%
  ggplot(aes(x=treatment, y=plant_region_area, fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill="none")+
  geom_text(data=Tk1, aes(label=cld1,x=treatment, y=quant),vjust=-5.3,hjust=-0.5, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"))+
  scale_shape_manual(values = c(16,17,15,18), labels=c("A1","A2","A3","A4"))+
  labs(x= "Treatment",
       y= "Rosette area (z-transformed)",
       title= "Rosette area (day 21)")+
  scale_x_discrete("Treatment", labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
 # scale_shape_discrete("Assay", labels=c("A1","A2","A3","A4"))+
  scale_y_continuous(limits=c(0,2.5))+
  theme(legend.position = "bottom")+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))
  #theme(plot.title = element_text(hjust=0.5))
F1
#ggsave("boxplot_plant_region_area.png",width=6, height=4) 
```


```{r,compact letter_prgs,echo=F, message=F, warning=F, results=F}

#plant_region_green_strength
anova_plant_region_green_strength<- aov(plant_region_green_strength~treatment, data=data_total)
tukey_plant_region_green_strength<-TukeyHSD(anova_plant_region_green_strength)
cld2<- multcompLetters4(anova_plant_region_green_strength, tukey_plant_region_green_strength)
print(cld2)

Tk2<- group_by(data_total,treatment) %>%
  summarise(mean=mean(plant_region_green_strength,na.rm=TRUE), quant=quantile(plant_region_green_strength, probs=0.75,na.rm=TRUE))%>%
  arrange(desc(mean))

cld2<- as.data.frame.list(cld2$treatment)
Tk2$cld2<-cld2$Letters
print(Tk2)

```

### Figure 2 \| Plant region green strength at day 21

\vspace{5mm}

```{r,plot_prgs,echo=F, message=F, warning=F}


#plant_region_green_strength

F2<- data_total %>%
  ggplot(aes(x=treatment, y=plant_region_green_strength, fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill="none")+
  geom_text(data=Tk2, aes(label=cld2,x=treatment, y=quant),vjust=-4.5,hjust=-0.8, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"))+
  scale_shape_manual(values = c(16,17,15,18), labels=c("A1","A2","A3","A4"))+
  labs(x= "Treatment",
       y= "Green strenght (z-transformed)",
       title= "Green strength of Rosette")+
  scale_x_discrete("Treatment", labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
  scale_y_continuous(limits=c(0,2))+
    theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=13))+
  theme(text=element_text(size=12, colour="black"))+
  theme(legend.position = "bottom")
  #theme(plot.title = element_text(hjust=0.5))
F2
#ggsave("boxplot_plant_region_green_strength.png",width=6, height=4) 

```


```{r,compact letter_prmal,echo=F, message=F, warning=F, results=F}

#plant_region_major_axis_length
anova_plant_region_major_axis_length<-aov(plant_region_major_axis_length~treatment, data=data_total)
tukey_plant_region_major_axis_length<-TukeyHSD(anova_plant_region_major_axis_length)
cld3<- multcompLetters4(anova_plant_region_major_axis_length,tukey_plant_region_major_axis_length)
print(cld3)

Tk3<- group_by(data_total,treatment) %>%
  summarise(mean=mean(plant_region_major_axis_length, na.rm=TRUE), quant=quantile(plant_region_major_axis_length, probs=0.75, na.rm=TRUE))%>%
  arrange(desc(mean))

cld3<- as.data.frame.list(cld3$treatment)
Tk3$cld3<-cld3$Letters
print(Tk3)
```

### Figure 3 \| Plant region major axis length at day 21

\vspace{5mm}

```{r,plot_prmal,echo=F, message=F, warning=F}

#plant_region_major_axis_length

F3<- data_total %>%
  ggplot(aes(x=treatment, y=plant_region_major_axis_length, fill=treatment))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(aes(shape=Assays), width = 0.3, alpha=0.6)+
  theme_classic()+
  guides(fill="none")+
  geom_text(data=Tk3, aes(label=cld3,x=treatment, y=quant),vjust=-4,hjust=-0.7, size=4)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"))+
  scale_shape_manual(values = c(16,17,15,18),labels=c("A1","A2","A3","A4"))+
  labs(x= "Treatment",
       y= "Major axis length (z-transformed)",
       title= "Major axis length of Rosette")+
  scale_x_discrete("Treatment", labels=c("HP C", " HP Syn9", "LP C", "LP Syn9"))+
  scale_y_continuous(limits=c(0,2))+
  theme(legend.position = "bottom")+
    theme(axis.title.x=element_blank())+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))
 # theme(plot.title = element_text(hjust=0.5))
F3
#ggsave("boxplot_plant_region_major_axis_length.png",width=6, height=4) 

```

```{r, plot_grid,echo=F, message=F, warning=F, fig.show='hide'}
fig <- plot_grid(F1 + labs(colour=""),
                F2 + labs(colour=""),
                F3 + labs(colour=""),
               labels=c("A","B","C"),
               align = "hv",
               rel_heights = c(1,1),
               nrow=2, ncol=2, label_size = 11)

legend<- get_legend(F1 +theme(legend.position = "bottom")+ labs(fill="treatment")+theme(legend.title = ))

title<- ggdraw()+draw_label("Biomass effects by SynCom9", fontface = "bold")


FIG<- plot_grid(title,fig,legend, nrow=3,ncol=1,rel_heights=c(.1, 1,.1 ))
FIG
#ggsave("FIG_Aradeepopsis.png",width=7, height=8) 


```


### Figure 4\| Plant region area
\vspace{5mm}

Plotting each treatment over time. Here we wanted to check at which time point our plants show the optimal growth. We will chose 3 features of interest (as many features show comparable characteristics in size, shape and colour): 

- for size: plant_region_area
- for shape: plant_region_minor_axis_length
- for colour: plant_region_green_strength


```{r ggplot smooth timepoint plant_region_area, echo=F, message=F, warning=F, fig.width=10, fig.height=4}
  
T1<-ggplot(data=data_foi_norm, aes(x=day, y=plant_region_area,color=treatment, group=treatment,outlier.shape = NA))+
  #geom_point(aes(col=treatment), size=1, pch=4)+
  geom_jitter(alpha=0.4)+
  geom_smooth(method="auto", se=FALSE, size=1)+
  scale_x_discrete()+
  theme_classic()+
  labs(x= "time point",
       y="mean area (z-transformed)",
       title="Rosette area over time")+
  scale_color_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"), name="Treatment",labels=c("HP C", "HP Syn9", "LP C", "LP Syn9") )+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))+
  theme(axis.ticks.x=element_blank(), legend.position = "right", axis.text.x = element_text(hjust=0), axis.title.x = element_blank())
T1
#ggsave("timepoint_plant_region_area.png", width = 6, height = 3) 

```

```{r timeseries, echo=F, message=F, warning=F, fig.width=10, fig.height=4}


Ta<-ggplot(data=data_foi_norm, aes(x=day, y=plant_region_area,color=treatment ,outlier.shape = NA), group=day)+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.6)+
  scale_x_discrete()+
  theme_classic()+
  labs(x= "time point",
       y="mean",
       title="Rosette area over time")+
  scale_color_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"), name="Treatment")+
  theme(axis.title.y = element_text(size=9))+
  theme(text=element_text(size=12, colour="black"))+
  theme(axis.ticks.x=element_blank(), legend.position = "right", axis.text.x = element_text(hjust=0), axis.title.x = element_blank())
Ta
#ggsave("timepoint_plant_region_area_Ta.png", width = 6, height = 3) 

```

```{r, plot_grid_with,echo=F, message=F, warning=F, fig.show='hide'}
fig <- plot_grid(F1 + labs(colour=""),
                 T1 + labs(colour=""),
               labels=c("A", "B"),
               align = "hv",
               rel_heights = c(1,1),
               rel_widths = c(4,8),
               nrow=1, ncol=2, label_size = 11)
fig

#ggsave("FIG_Aradeepopsis_tw.png",width=10, height=4) 


```

### Figure 5\| Plant region area major axis length

\vspace{5mm}

```{r ggplot smooth timepoint plant_region_major_axis_length, echo=F, message=F, warning=F, fig.width=10, fig.height=4}

ggplot(data=data_foi_norm, aes(x=day, y=plant_region_major_axis_length, colour=treatment, group=treatment,outlier.shape = NA))+
  #geom_point(aes(col=treatment), size=1, pch=4)+
  geom_jitter(alpha=0.4)+
  geom_smooth (method="auto", se=FALSE, size=1)+
  scale_x_discrete()+
  theme_classic()+
  labs(x= "time point",
       y="mean",
       title="Plant region major axis length")+
  theme(plot.title = element_text(hjust=0.5))+
  scale_color_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue","goldenrod1", "coral"), name="Treatment")
#ggsave("timepoint_plant_region_major_axis_length.png", width = 6, height = 3) 
```

### Figure 6\| Plant region green strength

\vspace{5mm}

```{r ggplot smooth timepoint plant_region_green_strength, echo=F, message=F, warning=F, fig.width=10, fig.height=4}

ggplot(data=data_foi_norm, aes(x=day, y=plant_region_green_strength, colour=treatment, group=treatment,outlier.shape = NA))+
  #geom_point(aes(col=treatment), size=1, pch=4)+
  geom_jitter(alpha=0.4)+
  geom_smooth (method="auto", se=FALSE, size=1)+
  scale_x_discrete()+
  theme_classic()+
  labs(x= "time point",
       y="mean",
       title="Plant region green strength")+
  theme(plot.title = element_text(hjust=0.5))+
  scale_color_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue","goldenrod1", "coral"),name="Treatment")
#ggsave("timepoint_plant_region_green_strength.png", width = 6, height = 3) 
```