---
title: "qPCR Analysis: Phosphor and nitrogen starvation response"
author: "Claudia Probst"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
    df_print: paged
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
geometry: margin=2cm
editor_options: 
  chunk_output_type: console
--- 
# Experimental Setup

We're investigating the responses of *Arabidopsis thaliana* to phosphate and nitrogen starvation responses. We suggest that in a low-P environment without SynCom9, plants suffer from phosphate starvation. 

\vspace{5mm}

Treatments: 

- low P control
- low P SynCom9
- high P control
- high P SynCom9
\vspace{5mm}

```{r setup, include=FALSE, echo=F, message=F, warning=F}
##clear the objects from memory ----
rm(list=ls())

# Load libraries ----
#library(pcr)
library(tidyverse)
library(readxl)
library(car)
library(multcompView)
library (cowplot)

# Load data ----
setwd("~/Desktop/MASTER Pflanzenwissenschaften/Experiments/qPCR/results")

qPCR_pl1 <- read_excel("qPCR_plate1.xlsx", 
                       sheet = "23-02-01-qPCRPlate1-Cla_compact", 
                       range = "A4:H100")

qPCR_pl2 <- read_excel("qPCR_plate2.xlsx", 
                       sheet = "23-02-01-qPCRPlate2-Cla_compact", 
                       range = "A4:H88")

meta <- read_excel("design.xlsx", sheet = "Tabelle1") 

meta1 <- meta %>% 
  filter(plate==1) %>% 
  select(Well, gene, sample_ID, sample_name, identification)

meta2 <- meta %>% 
  filter(plate==2) %>% 
  select(Well, gene, sample_ID, sample_name, identification)

# Tidy the data plate 1 ----

qPCR1 <- qPCR_pl1 %>% 
  separate(name, c("well", "sample_nr"), sep = "_") %>% 
  left_join(meta1, by = c("well" = "Well")) %>% 
  unique()

# Tidy the data plate 2 ----

qPCR2 <- qPCR_pl2 %>% 
  separate(name, c("well", "sample_nr"), sep = "_") %>% 
  left_join(meta2, by = c("well" = "Well")) %>% 
  unique()

rm(meta)  
rm(meta1)
rm(meta2)
rm(qPCR_pl1)
rm(qPCR_pl2)

#combine all data ----
qPCR <- rbind(qPCR1, qPCR2)

rm(qPCR1)
rm(qPCR2)

```

```{r AT4 gene, echo=F, message=F, warning=F }

#checking for normal distribution of AT4 gene revealed no normal distribution
LPS<- subset(qPCR, Amplicon=="AT4_SYN_lowP")
HPS<- subset(qPCR, Amplicon=="AT4_SYN_highP")
HPC<- subset(qPCR, Amplicon=="AT4_CONTROL_highP")
LPC<- subset(qPCR, Amplicon=="AT4_CONTROL_highP")

#log transformation because data are not normally distributed

LPC$Cq_new <- log(LPC$Cq)
LPS$Cq_new <- log(LPS$Cq)
HPS$Cq_new <- log(HPS$Cq)
HPC$Cq_new <- log(HPC$Cq)

par(mfrow=c(3:2))
hist(LPC$Cq_new)
qqplot<-scale(LPC$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPC$Cq_new)

hist(LPS$Cq_new)
qqplot<-scale(LPS$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPS$Cq_new)

hist(HPS$Cq_new)
qqplot<-scale(HPS$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPS$Cq_new)

par(mfrow=c(3:2))
hist(HPC$Cq_new)
qqplot<-scale(HPC$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPC$Cq_new)

#Still, data is not normally distributed
```

```{r CEPD2 gene, echo=F, message=F, warning=F }

#checking for normal distribution of CEPD2 gene revealed no normal distribution

LPS<- subset(qPCR, Amplicon=="CEPD2_SYN_lowP")
HPS<- subset(qPCR, Amplicon=="CEPD2_SYN_highP")
HPC<- subset(qPCR, Amplicon=="CEPD2_CONTROL_highP")
LPC<- subset(qPCR, Amplicon=="CEPD2_CONTROL_lowP")

#log transformation because data are not normally distributed

LPC$Cq_new <- log(LPC$Cq)
LPS$Cq_new <- log(LPS$Cq)
HPS$Cq_new <- log(HPS$Cq)
HPC$Cq_new <- log(HPC$Cq)

par(mfrow=c(3:2))
hist(LPC$Cq_new)
qqplot<-scale(LPC$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPC$Cq_new)

hist(LPS$Cq_new)
qqplot<-scale(LPS$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(LPS$Cq_new)

hist(HPS$Cq_new)
qqplot<-scale(HPS$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPS$Cq_new)

par(mfrow=c(3:2))
hist(HPC$Cq_new)
qqplot<-scale(HPC$Cq_new)
qqnorm(qqplot)
qqline(qqplot)
shapiro.test(HPC$Cq_new)

#Still, data is not normally distributed
```


```{mean of replicates, echo=F, message=F, warning=F}

# Calculate mean of replicates per sample

ct <- qPCR %>% 
  select(well,identification, sample_nr, sample_name, gene, Cq) %>% 
  filter(!identification %in% c("5","25","31","32","33","34","38","39","40","41","42","43","44","45","46","47","48","100","101","102","104","107","112","113","114")) %>%
  pivot_wider(names_from = gene, values_from = c(Cq)) %>% 
  group_by(sample_nr, sample_name) %>% 
  summarise(av_AT4=mean(AT4, na.rm=TRUE),
            av_CEPD2=mean(CEPD2, na.rm = TRUE),
            av_reference=mean(reference, na.rm = TRUE))

E <- qPCR %>% 
  group_by(gene) %>% 
  summarise(E=mean(mean_PCR_eff)) %>% 
  pivot_wider(names_from = gene, values_from = E)

Ct_control <- ct %>% 
  filter(sample_name=="lowP_control") %>% 
  group_by(sample_name) %>% 
  summarise(Ct_control_AT4=mean(av_AT4, na.rm=TRUE),
            Ct_control_CEPD2=mean(av_CEPD2,na.rm=TRUE),
            Ct_control_reference=mean(av_reference,na.rm=TRUE))


```

```{delta-delta Ct method, echo=F, message=F, warning=F} 

deltadelta <- ct %>% 
  mutate(delta_AT4=Ct_control$Ct_control_AT4 - av_AT4) %>% 
  mutate(delta_CEPD2=Ct_control$Ct_control_CEPD2 - av_CEPD2) %>%
  mutate(delta_reference=Ct_control$Ct_control_reference - av_reference) %>%
  mutate(norm_expr_AT4=(E$AT4^delta_AT4)/(E$reference^delta_reference)) %>%
  mutate(norm_expr_CEPD2=(E$CEPD2^delta_CEPD2)/(E$reference^delta_reference))

```

```{r Kruksal test,echo=F, message=F, warning=F}
# AT4 ----

Kruk1<-kruskal.test(norm_expr_AT4~sample_name, data=deltadelta)
Kruk1


Kruk2<-kruskal.test(norm_expr_CEPD2~sample_name, data=deltadelta)
Kruk2


```

```{r}
library(dunn.test)
library(rstatix)

dt1<-dunn.test(deltadelta$norm_expr_AT4,deltadelta$sample_name,method="bonferroni", kw=TRUE, wrap=TRUE)
dt2<-dunn.test(deltadelta$norm_expr_CEPD2,deltadelta$sample_name,method="bonferroni", kw=TRUE, wrap=TRUE)

```

```{r figure, echo=F, message=F, warning=F}
#my.comparison<-list(c("highP_control", "lowP_control"), c("lowP_control", "lowP_Syn9"),c("lowP_control", "highP_Syn9"),c("lowP_Syn9", "highP_Syn9"),c("highP_control", "highP_Syn9" ))

library(ggpubr)
library(multcompView)
library(rcompanion)

letters1 <- cldList(P.adjusted ~ comparisons, data=dt1, threshold=0.05)
letters2 <- cldList(P.adjusted ~ comparisons, data=dt2, threshold=0.05)


fig1<- deltadelta %>% 
  select(norm_expr_AT4, sample_nr, sample_name) %>% 
  pivot_longer(c(norm_expr_AT4), names_to = "gene", values_to = "norm_rel_expr") %>% 
  ggplot(aes(x=sample_name, y=norm_rel_expr, fill=sample_name))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(width = 0.3, alpha=0.6)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"))+
  guides(fill="none")+
  theme_classic()+
  annotate("text", x=4, y=4, label="")+
  labs(x= "Treatment",
       y= "relative expression",
       title= "Phosphate Starvation Response (AT4)")+
  annotate("text", x=1.1, y=0.5, label="ab")+
  annotate("text", x=2.1, y=0.5, label="a")+
  annotate("text", x=3.1, y=2.8, label="ab")+
  annotate("text", x=4.1, y=2.8, label="b")+
 #stat_compare_means(comparisons=my.comparison,method="wilcox.test")+
  theme(plot.title = element_text(hjust=0.5))+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))+
  scale_x_discrete(labels=c("HP C", "HP Syn8", "LP C", "LP Syn8"))
  #facet_wrap(~gene, scales = "free")
fig1

fig2<- deltadelta %>% 
  select(norm_expr_CEPD2, sample_nr, sample_name) %>% 
  pivot_longer(c(norm_expr_CEPD2), names_to = "gene", values_to = "norm_rel_expr") %>% 
  ggplot(aes(x=sample_name, y=norm_rel_expr, fill=sample_name))+
  geom_boxplot(outlier.shape = NA,alpha=0.8,width = 0.4)+
  geom_jitter(width = 0.3, alpha=0.6)+
  scale_fill_manual(values = c("darkolivegreen3","#339966","lightskyblue1","cornflowerblue"))+
  guides(fill="none")+
  theme_classic()+
  labs(x= "Treatment",
       y= "relative expression",
       title= "Nitogen Starvation Response (CEPD2)")+
  annotate("text", x=1.1, y=7, label="a")+
  annotate("text", x=2.1, y=7, label="ab")+
  annotate("text", x=3.1, y=7, label="a")+
  annotate("text", x=4.1, y=60, label="b")+
  theme(plot.title = element_text(hjust=0.5))+
  theme(axis.title.y = element_text(size=12))+
  theme(text=element_text(size=12, colour="black"))+
  scale_x_discrete(labels=c("HP C", "HP Syn8", "LP C", "LP Syn8"))
  #facet_wrap(~gene, scales = "free")
fig2

#combine the two plots ----

fig <-cowplot::plot_grid(fig1,
                          fig2 + labs(colour=""),
                          labels=c("A","B"),
                          align = "hv",
                          rel_heights = c(6,6),
                          nrow=1, ncol=2)
fig


# save plots ----
ggsave("qPCR_fig2.png", width = 8, height = 4)


```

# Analysis using N0 value directly (Henrys method) ----
qPCR1_N0 <- qPCR %>% 
  group_by(sample_nr, sample_name, gene) %>% 
  summarise(N0Av=mean(N0)) %>% 
  pivot_wider(names_from = gene, values_from = N0Av) %>% 
  mutate(AT4norm=AT4/reference,
         CEPD2norm=CEPD2/reference)

plot2 <- qPCR1_N0 %>% 
  select(AT4norm, CEPD2norm, sample_nr, sample_name) %>% 
  pivot_longer(c(AT4norm, CEPD2norm), names_to = "gene", values_to = "norm_rel_expr") %>% 
  ggplot(aes(x=sample_name, y=norm_rel_expr))+
  geom_boxplot() +
  facet_wrap(~gene, scales = "free")


# This does the same! The only difference here is that the expression is not yet calculated relative to the control

